# Exercise 1

The time it takes a student to complete a TOLC-I University orientation and evaluation test follows a density function of the form:

$$ f(t) = \begin{cases} c(t−1)(2−t) & 1< t <2 \\ 0 & \mbox{othervise} \end{cases} $$
where $t$ is the time in hours.

1. using the `integrate()` $R$ function, determine the constant $c$ (and verify it analytically)
1. write the set of four $R$ functions and plot the pdf and cdf, respectively
1. evaluate the probability that the student will finish the aptitude test in:
    - more than $75$ minutes;
    - between $90$ and $120$ minutes.


## Solution

**1.** The constant $c$ needs to be set as normalization constant: in fact, in order for $f$ to be a $pdf$, the area below the curve must be equal to $1$. Hence:

$$ \frac{1}{c} = \int^2_1 (t-1)(2-t)dt = \int^2_1 (3t - t^2 -2) dt = \left[ \frac{3}{2}t^2 - \frac{t^3}{3} - 2t \right]^2_1 = 6 - \frac{8}{3} - 4 - \frac{3}{2} + \frac{1}{3} + 2 = \frac{1}{6} \implies c=6$$

The result obtained above is derived again by integrating $f$ with the `integrate` $R$ function in the cell below. 

```{r}
f <- function(t){
    y <- ifelse ( (t<2 & t>1),
                  (t-1)*(2-t),
                  0)    
    return(y)
}
f<-Vectorize(f)

area <- integrate(f, 1, 2)

(c<-1/area$value)
```
**2.** The functions in the cell below are:
 
- the probability distribution function $f$;
- the cumulative distribution function $F$, defined as $F(x)=\int^x_{-\infty}f(t) dt$;
- the quantile function $q(y) = F^{-1}(y)$;
- a function to randomly extract $n$ samples with distribution $f$ named $rf$.

Some of them are derived with different methods that yield the same results.

```{r}
#install.packages('GoFKernel')
library('GoFKernel')

#-------------PDF-------------

#the value of c = 6 is derived above
pdf <- function(t){
    y <- ifelse ( (t<2 & t>1),
                  c*(t-1)*(2-t),
                  0   
                )
    return(y)
}
pdf <- Vectorize(pdf)

#-------------CDF-------------

# Derived by manually integrating, obtained analytically
cdf <- function(t){
    y <- ifelse ( (t<2 & t>1),
                   9*t^2 - 2*t^3 - 12*t + 5,
                 ifelse ( (t>=2),
                           1,
                           0)
                 )
    return(y)
} 
cdf <- Vectorize(cdf)

#Derived via its definition and the integrate R function
cdf_i <- function(t){
    y <- ifelse ( (t<2 & t>1),
                   integrate(pdf, 1, t)$value,
                 ifelse ( (t>=2),
                           1,
                           0)
                 )
    return(y)
} 
cdf_i <- Vectorize(cdf_i)

#-------------QUANTILE FUNCTION-------------

#Derived by inverse function of GoFKernel package
q <- Vectorize(inverse(cdf, 1, 2))       #upper and lower boundaries of cdf domain

#Derived by "manual" method and root finding function uniroot
inv <- function (f, lower, upper) {
   function (y){
     uniroot(function(x){f(x)-y}, lower = lower, upper = upper)$root 
   } 
}
qf <- Vectorize(inv(cdf, 1, 2))          #upper and lower boundaries of cdf domain

#-------------RANDOM EXTRACTION OF n SAMPLES-------------

#Inverse trasform method
rf <- function(n, cdf, lower, upper) {
    us <- runif(n)
    rnd <- Vectorize(inverse(cdf, lower = lower, upper = upper)) 
    return (rnd(us))
}

```

```{r}
options(repr.plot.width=18, repr.plot.height=12)  #to set graph size
par(mfrow=c(2,3),  mar=c(5, 5, 4, 2))             #graphs in the same row

x <- seq(1, 2, 0.04)
p <- seq(0, 1, 0.04)

plot(x, pdf(x), lwd=2.5, pch=4, col="red",
     main="Probability distribution function", cex.main=2,
     xlab="x [hours]", ylab="f(x)", cex.lab=2, las=1)
lines(x, pdf(x), lwd=2.5, lty="dashed", col="darkred")

plot(x, cdf(x), lwd=2.5, pch=4, col="blue",
     main="Cumulative distribution function, analytical method", cex.main=2,
     xlab="x [hours]", ylab="F(x)", cex.lab=2, las=1)
lines(x, cdf(x), lwd=2.5, lty="dashed", col="dodgerblue2")

plot(x, cdf_i(x), lwd=2.5, pch=4, col="orange",
     main="Cumulative distribution function, integrate method", cex.main=2,
     xlab="x [hours]", ylab="F(x)", cex.lab=2, las=1)
lines(x, cdf_i(x), lwd=2.5, lty="dashed", col="gold")

plot(p, q(p), lwd=2.5, pch=4, col="forestgreen",
     main="Quantile function", cex.main=2,
     xlab="p", ylab="q(p)", cex.lab=2, las=1)
lines(p, q(p), lwd=2.5, lty="dashed", col="darkolivegreen1")

plot(p, qf(p), lwd=2.5, pch=4, col="darkorchid",
     main="Quantile function, manual method", cex.main=2,
     xlab="p", ylab="q(p)", cex.lab=2, las=1)
lines(p, qf(p), lwd=2.5, lty="dashed", col="plum")

h <- hist(rf(10000, cdf, 1, 2), breaks= 15, col= "aquamarine", 
          main="10000 Random samples from pdf", cex.main=2,
          xlab="x [hours]", ylab="Counts", cex.lab=2, las=1)
lines(x, 10000*(h$breaks[2]-h$breaks[1])*pdf(x), lwd=2.5, lty="dashed", col="grey")

```

**3.** To evaluate the probability that the student will finish the aptitude test in a given time (more than $75$ minutes, between $90$ and $120$ minutes), there are two equivalent methods:

- by integrating the pdf with the given boundaries;
- by using the cdf.

It can be seen that both methods leads to the same numerical results: more than $84\%$ of students take the test in more than $75$ minutes, while for half of them ($50\%$) it takes from $90$ minutes to $2$ hours.

```{r}
#simple function to convert from minutes to hours in order to fit the pdf x-domain
mins_to_h <- function(min){
    h <- min/60 
    return(h)
}

p_75 <- integrate(pdf, mins_to_h(75), 2)
p_90_120 <- integrate(pdf, mins_to_h(90), mins_to_h(120))

p_75_c <- 1 - cdf(mins_to_h(75))
p_90_120_c <- cdf(mins_to_h(120))- cdf(mins_to_h(90))

print(paste("P(x>1.25)_integrate=",p_75$value))
print(paste("P(x>1.25)_cdf=",p_75_c))
print(paste("P(1.5<x<2)_integrate=",p_90_120$value))
print(paste("P(1.5<x<2)_cdf=",p_90_120_c))
```

# Exercise 2

The lifetime of tires sold by an used tires shop is $10^4·x$ km, where $x$ is a random variable following the distribution function:

$$ f(x) = \begin{cases} \frac{2}{x^2} & 1< x <2 \\ 0 & \mbox{othervise} \end{cases} $$

1. write the set of four $R$ functions and plot the pdf and cdf, respectively
1. determine the probability that tires will last less than $15000$ km
1. sample $3000$ random variables from the distribution and determine the mean value and the variance, using the expression $Var[x]=E[x^2]- E^2[x]$

## Solution

**1.** The functions are defined exactly as **Exercise 1**, but this time only one method to derive them is chosen. The plots are displayed below.

```{r}
#-------------PDF-------------

pdf <- function(x){
    y <- ifelse ( (x<=2 & x>=1),
                   2/(x^2),
                   0)
    return(y)
}
pdf <- Vectorize(pdf)

#-------------CDF-------------

#Derived via its definition and the integrate R function
cdf <- function(x){
    y <- ifelse ( (x<=2 & x>=1),
                   integrate(pdf, 1, x)$value,
                 ifelse ( (x>2),
                           1,
                           0)
                 )
    return(y)
} 
cdf <- Vectorize(cdf_i)

#-------------QUANTILE FUNCTION-------------

#Derived by inverse function of GoFKernel package
q <- Vectorize(inverse(cdf, 1, 2))       #upper and lower boundaries of cdf domain

#-------------RANDOM EXTRACTION OF n SAMPLES-------------

#Inverse trasform method
rf <- function(n, cdf, lower, upper) {
    us <- runif(n)
    rnd <- Vectorize(inverse(cdf, lower = lower, upper = upper)) 
    return (rnd(us))
}
```

```{r}
options(repr.plot.width=14, repr.plot.height=6)  #to set graph size
par(mfrow=c(1,2))#,  mar=c(5, 5, 4, 2))           #graphs in the same row

x <- seq(1, 2, 0.04)

plot(x, pdf(x), lwd=2.5, pch=4, col="red",
     main="Probability distribution function", cex.main=1.5,
     xlab=expression(paste("x [", 10^4, " km]")), ylab="f(x)", cex.lab=1.5, las=1)
lines(x, pdf(x), lwd=2.5, lty="dashed", col="darkred")

plot(x, cdf(x), lwd=2.5, pch=4, col="blue",
     main="Cumulative distribution function", cex.main=1.5,
     xlab=expression(paste("x [", 10^4, " km]")), ylab="F(x)", cex.lab=1.5, las=1)
lines(x, cdf(x), lwd=2.5, lty="dashed", col="dodgerblue2")

```

**2.** The probability that tires will last less than $15000$ km is given by the integral of the $pdf$ for $x<1.5$, or the $cdf$ with $x=1.5 \implies 66.67\%$.

```{r}
prob <- cdf(1.5)
print(paste("Probability of tires lasting less than 15000 km (cdf): ", prob))

p_pdf <- integrate(pdf, 1, 1.5)
print(paste("Probability of tires lasting less than 15000 km (pdf): ", p_pdf$value))

```

**3.** By sampling $3000$ random variables from the distribution, it is possible to determine its mean value and variance, knowing that:

$$E[x]= \int^2_1 x f(x) dx$$

and $$Var[x]=E[x^2]- E^2[x].$$

Two estimates of the average value are obtained:

- from the simulated data, by averaging them with the `mean` function;
- from the $pdf$, with the definition above.

The cell below shows the "theoretical" value derived with the `integrate` function - with its associated error given by the function (which has not a statistical meaning), and the "experimental" value derived from the simulated data, whose error is the squared root of the variance $Var[x]$ of the distribution.

```{r}
sim_3000 <- rf(3000, cdf, 1, 2)

x <- seq(1,2,1/3000)

avg_exp <- mean(sim_3000)
e_x <- integrate(function(x){x * pdf(x)}, 1, 2)
e_x2 <- integrate(function(x){x^2 * pdf(x)}, 1, 2)

var <-  e_x2$value - e_x$value^2
var
print(paste("Theoretical mean:", e_x$value, "+/-", e_x$abs.error))
print(paste("Experimental mean:", avg_exp, "+/-", sqrt(var)))

```

<!-- #region -->
# Exercise 3

Markov’s inequality represents an upper boud to probability distributions:

$$ P(X \geq k) \leq \frac{E[X]}{k} \mbox{ for } k >0 $$

Having defined a function

$$ G(k) = 1−F(k) \equiv P(X \geq k) $$

plot $G(k)$ and the Markov’s upper bound for:

1. the exponential distribution function $Exp(\lambda= 1)$;
1. the uniform distribution function $U(3,5)$;
1. the binomial distribution function $Bin(n= 1,p= 1/2)$;
1. the Poisson distribution function $Pois(\lambda= 1/2)$.


## Solution
<!-- #endregion -->

```{r}
#-----Definition of the left-hand-side of Markov's inequality functions-----

g_exp  <- function(k) {1- pexp(k, 1)}         #exponential distribution
g_unif <- function(k) {1- punif(k, 3, 5)}     #uniform distribution
g_bin  <- function(k) {1- pbinom(k, 1, 0.5)}  #binomial distribution
g_pois <- function(k) {1- ppois(k, 0.5)}      #Poisson distribution 

#-----Definition of the expected values for the given distributions-----

E_exp  <- 1     #1/lambda
E_unif <- 4     #(a+b)/2
E_bin  <- 1/2   #n*p
E_pois <- 1/2   #lambda

#-----Definition of the right-hand-side of M's inequality-----

markov  <- function(k, E) {E/k}
```

```{r}
options(repr.plot.width=14, repr.plot.height=12)  #to set graph size
par(mfrow= c(2,2), mar=c(5, 5, 4, 2))

x <-seq(0, 10, 0.5)

plot(x, g_exp(x), lwd=2.5, pch=4, col="red",
     main=expression(paste("1. Exponential distribution, Exp(", lambda,"=1)")), cex.main=2,
     xlab="k", ylab=expression("P[X">="k]"), cex.lab=1.5, las=1)
lines(x, g_exp(x), lwd=2.5, lty="dashed", col="darkred")
curve(markov(x, E_exp), from=0, to=10, lwd=2, lty="dotdash", col="black", add=TRUE)
legend(5, 1.1, legend=c("Markov", expression("P[X">="k]")),
       col=c("black", "darkred"), lty=c("dotdash", "dashed"), 
       bty='n', cex = 1.4 )

plot(x, g_unif(x), lwd=2.5, pch=4, col="blue",
     main=expression(paste("2. Uniform distribution, U(3,5)")), cex.main=2,
     xlab="k", ylab=expression("P[X">="k]"), cex.lab=1.5, las=1)
lines(x, g_unif(x), lwd=2.5, lty="dashed", col="dodgerblue2")
curve(markov(x, E_unif), from=0, to=10, lwd=2, lty="dotdash", col="black", add=TRUE)
legend(5, 1.1, legend=c("Markov", expression("P[X">="k]")),
       col=c("black", "dodgerblue2"), lty=c("dotdash", "dashed"), 
       bty='n', cex = 1.4 )

plot(x, g_bin(x), lwd=2.5, pch=4, col="forestgreen",
     main=expression(paste("3. Binomial distribution, Bin(n=1, p=1/2)")), cex.main=2,
     xlab="k", ylab=expression("P[X">="k]"), cex.lab=1.5, las=1)
lines(x, g_bin(x), type='s', lwd=2.5, lty="dashed", col="darkolivegreen1")
curve(markov(x, E_bin), from=0, to=10, lwd=2, lty="dotdash", col="black", add=TRUE)
legend(5, 0.55, legend=c("Markov", expression("P[X">="k]")),
       col=c("black", "forestgreen"), lty=c("dotdash", "dashed"), 
       bty='n', cex = 1.4 )

plot(x, g_pois(x), lwd=2.5, pch=4, col="darkorchid",
     main=expression(paste("4. Poisson distribution, Pois(", lambda,"=1/2)")), cex.main=2,
     xlab="k", ylab=expression("P[X">="k]"), cex.lab=1.5, las=1)
lines(x, g_pois(x), type='s', lwd=2.5, lty="dashed", col="plum")
curve(markov(x, E_pois), from=0, to=10, lwd=2, lty="dotdash", col="black", add=TRUE)
legend(5, 0.43, legend=c("Markov", expression("P[X">="k]")),
       col=c("black", "darkorchid"), lty=c("dotdash", "dashed"), 
       bty='n', cex = 1.4 )

```

<!-- #region -->
# Exercise 4

Chebyshev’s inequality tell us that:

$$P(|X−\mu|≥k\sigma)≤\frac{1}{k^2}$$

which can also be written as:

$$P(|X−\mu|< k\sigma)≥1−\frac{1}{k^2}$$

Use  $R$  to  show,  with  a  plot,  that  Chebyshev’s  inequality  is  is  an  upper  bound  to  the  following distributions: 

1. a normal distribution, $N(\mu= 3,\sigma= 5)$;
1. an exponential distribution, $Exp(\lambda= 1)$;
1. a uniform distribution, $U(1−\sqrt{2},1 +\sqrt{2})$;
1. a Poisson distribution function, $Pois(\lambda= 1/3)$.


## Solution


<!-- #endregion -->

```{r}
mu_n      <- 3         #average of normal distribution
sigma_n   <- 5         #sqrt(variance) of normal distribution
mu_exp    <- 1         #average = 1/lambda
sigma_exp <- 1         #sqrt(variance) =1/lambda
mu_u      <- 1         #mean = (a+b)/2
sigma_u   <- sqrt(2/3) #sqrt(variance) = (b-a)/sqrt(12) 
lambda    <- 1/3       #mean Pois = var Pois = lambda 

p_n  <- function(k, mu, sigma){
            pnorm(mu_n+k*sigma_n, mu_n, sigma_n)-pnorm(mu_n-k*sigma_n, mu_n, sigma_n)
        }
p_e  <- function(k, mu, sigma){
            pexp(mu_exp+k*sigma_exp, mu_exp)-pexp(mu_exp-k*sigma_exp, mu_exp)
        }
p_u  <- function(k, mu, sigma){
            punif(mu_u+k*sigma_u, 1-sqrt(2), 1+sqrt(2))-punif(mu_u-k*sigma_u, 1-sqrt(2), 1+sqrt(2))
        }
p_p  <- function(k, mu, sigma){
            ppois(lambda+k*sqrt(lambda), lambda)-ppois(lambda-k*sqrt(lambda), lambda)
        }

limit  <- function(k) {1- (1/k)^2}
```

```{r}
options(repr.plot.width=14, repr.plot.height=12)  #to set graph size
par(mfrow= c(2,2), mar=c(5, 5, 4, 2))

k <- seq(0,4,0.1)

plot(k, p_n(k), lwd=2.5, pch=4, col="red",
     main=expression(paste("1. Normal distribution, N(",mu,"=3,",sigma,"=5)")), cex.main=2,
     xlab="k", ylab=expression(paste("P[|X-",mu,"|<k",sigma,"]")), 
     cex.lab=1.5, las=1)
lines(k, p_n(k), lwd=2.5, lty="dashed", col="darkred")
curve(limit(x), from=0, to=4, lwd=2, lty="dotdash", col="black", add=TRUE)
legend(2, 0.3, legend=c("Chebyshev", expression(paste("P[|X-",mu,"|<k",sigma,"]"))),
       col=c("black", "darkred"), lty=c("dotdash", "dashed"), 
       bty='n', cex = 1.4 )

plot(k, p_e(k), lwd=2.5, pch=4, col="blue",
     main=expression(paste("2. Exponential distribution, Exp(", lambda,"=1)")), cex.main=2,
     xlab="k", ylab=expression(paste("P[|X-",mu,"|<k",sigma,"]")), 
     cex.lab=1.5, las=1)
lines(k, p_e(k), lwd=2.5, lty="dashed", col="dodgerblue2")
curve(limit(x), from=0, to=4, lwd=2, lty="dotdash", col="black", add=TRUE)
legend(2, 0.3, legend=c("Chebyshev", expression(paste("P[|X-",mu,"|<k",sigma,"]"))),
       col=c("black", "dodgerblue2"), lty=c("dotdash", "dashed"), 
       bty='n', cex = 1.4 )

plot(k, p_u(k), lwd=2.5, pch=4, col="forestgreen",
     main=expression(paste("3. Uniform distribution, U(1-",sqrt(2),",1+",sqrt(2),")")), cex.main=2,
     xlab="k", ylab=expression(paste("P[|X-",mu,"|<k",sigma,"]")), 
     cex.lab=1.5, las=1)
lines(k, p_u(k), lwd=2.5, lty="dashed", col="darkolivegreen1")
curve(limit(x), from=0, to=4, lwd=2, lty="dotdash", col="black", add=TRUE)
legend(2, 0.3, legend=c("Chebyshev", expression(paste("P[|X-",mu,"|<k",sigma,"]"))),
       col=c("black", "darkolivegreen1"), lty=c("dotdash", "dashed"), 
       bty='n', cex = 1.4 )

plot(k, p_p(k), xlim=c(0,4), ylim=c(0,1), lwd=2.5, pch=4, col="darkorchid",
     main=expression(paste("4. Poisson distribution, Pois(", lambda,"=1/3)")), cex.main=2,
     xlab="k", ylab=expression(paste("P[|X-",mu,"|<k",sigma,"]")), 
     cex.lab=1.5, las=1)
lines(k, p_p(k), lwd=2.5,  type='s', lty="dashed", col="plum")
curve(limit(x), from=0, to=4, lwd=2, lty="dotdash", col="black", add=TRUE)
legend(2, 0.3, legend=c("Chebyshev", expression(paste("P[|X-",mu,"|<k",sigma,"]"))),
       col=c("black", "plum"), lty=c("dotdash", "dashed"), 
       bty='n', cex = 1.4 )

```
