# Exercise 1

The triangular distribution, in the interval $(a,b)$, is given by the following:

$$ f(x) = \begin{cases} \frac{2(x−a)}{(b−a)(c−a)} & a \leq x < c     \\ 
                        \frac{2(b−x)}{(b−a)(b−c)} & c \leq x \leq b  \\
                        0                & \mbox{othervise} \\    
          \end{cases} $$
where $c \in (a,b)$.

1. plot the function, given the interval $(a,b)$...
1. ... and write an algorithm to generate random numbers from the triangular distribution
1. generate $10^4$ random number from the distribution, show them in an histogram and superimpose the analytical curve


## Solution

**1.**  The following function $f$ allow us to calculate the triangular distribution given the interval $(a,b)$ and the mode $c$. To properly works, $c$ must be inside the $(a,b)$ interval and $a$ must be strictly smaller than $b$.


```{r}
f <- function(x, a, b, c){
    y <- ifelse ( (x<c & x>=a),
                   2*(x−a)/((b−a)*(c−a)),
                   ifelse ( (x<=b & x>=c),
                             2*(b−x)/((b−a)*(b−c)),
                             0)
                )
    return(y)
}
f <- Vectorize(f)
```
In the following plots, the triangular distribution is represented with three different $(a,b)$ intervals and $c$s.

```{r}
options(repr.plot.width=18, repr.plot.height=6)  #to set graph size
par(mfrow=c(1,3),  mar=c(5, 5, 4, 2))            #graphs in the same row

x <- seq(-6, 6, 0.05)

plot(x, f(x, -4, 4, 1), xlim=c(-6,6),
     lwd=2.5, pch=4, col="red",
     main="(a, b) = (-4, 4); c = 1", cex.main=2,
     xlab="x", ylab="f(x)", cex.lab=2, las=1)
lines(x, f(x, -4, 4, 1), lwd=2.5, lty="dashed", col="darkred")

plot(x, f(x, -1, 2.5, 1), xlim=c(-6,6),
     lwd=2.5, pch=4, col="blue",
     main="(a, b) = (-1, 2.5); c = 1", cex.main=2,
     xlab="x", ylab="f(x)", cex.lab=2, las=1)
lines(x, f(x, -1, 2.5, 1), lwd=2.5, lty="dashed", col="dodgerblue2")

plot(x, f(x, 0, 5, 4), xlim=c(-6,6),
     lwd=2.5, pch=4, col="orange",
     main="(a, b) = (0, 5); c = 4", cex.main=2,
     xlab="x", ylab="f(x)", cex.lab=2, las=1)
lines(x, f(x, 0, 5, 4), lwd=2.5, lty="dashed", col="gold")


```

**2.** To generate random numbers from the triangular distribution, two methods are implemented.

#### Acceptance/Rejection method 
In the `gen_samples_f` function, the Acceptance/Rejection method is coded: two vectors of length $n$ are generated with uniform distribution, one by drawing samples $x$ in the $(a, b)$ interval (`xs`), the other one sampling probabilities $u$ between $0$ and $1$ (`us`). The samples $x$ are accepted if $f(x)>u\cdot c$. The algorithm stops when exactly `n` samples are accepted.

#### Inverse trasform method
The first step to implement the Inverse trasform method is to calculate the cumulative density function $F$ by integrating $f$:

$$ F(x)= \int_{-\infty}^x f(t) dt = \left\{\begin{array}{ll}
\frac{(x-a)^{2}}{(b-a)(c-a)} & a \leq x<c \\
1-\frac{(x-b)^{2}}{(b-a)(b-c)} & c \leq x \leq b
\end{array}\right. $$

 Then, the quantile function $q$ is derived by inverting $F$:

$$q(u) = F^{-1}(u)=\left\{\begin{array}{ll}
a+\sqrt{(b-a)(c-a) u} & 0 \leq u<\frac{c-a}{b-a} \\
b-\sqrt{(b-a)(b-c)(1-u)} & \frac{c-a}{b-a} \leq u \leq 1
\end{array}\right.$$

Finally, $n$ samples $p$ in $(0,1)$ are drawn uniformly, and applying to them $F^{-1}$, the results obtained are $n$ $f$-distributed samples.

```{r}
gen_samples_f <- Vectorize(function(n, a, b, c){
                y <- NULL
                while(length(y)<n){
                    xs <- runif(n, a, b)
                    us <- runif(n, 0, 1)

                    ys <- ifelse(us * c < f(xs, a, b, c),
                                xs,
                                NA)

                    ys <- na.omit(ys)
                    y <- c(y, ys)
                    }
                return(y)
                })

```

```{r}
inv_transform <- Vectorize(function(n, a, b, c){
                ps <- runif(n, 0, 1)
                xs <- ifelse( (ps <(c−a)/(b−a)),
                              a + sqrt((b-a)*(c-a)*ps),
                              b - sqrt((b-a)*(b-c)*(1-ps))
                            )
                return(xs)
                })

```

**3.** After generating $10^4$ random number from the distribution, two histograms are plotted, one for each method implemented. 
One of them is a density histogram, normalized to the number of samples; the other one is a frequency histogram, hence $f$ in this case is multiplied by the actual area of the histogram ($w\cdot n$, with $w$ bin width and $n$ number of samples).

$a$, $b$ and $c$ are arbitrarily fixed to $-4,4,1$.

```{r}
a <- -4 
b <- 4
c <- 1

x <- seq(-6, 6, 0.5)
n <- 10000

options(repr.plot.width=16, repr.plot.height=8)  #to set graph size
par(mfrow=c(1,2),  mar=c(5, 5, 4, 2))            #graphs in the same row

hist(gen_samples_f(n, a, b, c), freq=FALSE,
     col="lightskyblue", xlab="x", xlim=c(-6,6), ylim=c(0,0.3),
     main="Acceptance/Rejection sampling, density histogram")
curve(f(x, a, b, c), col="royalblue4", lt="dashed", lw=3, add=TRUE)

h <- hist(inv_transform(n, a, b, c), freq=TRUE,
     col="mistyrose", xlab="x", xlim=c(-6,6), 
     main="Inverse trasform sampling, frequency histogram")
lines(x, n*(h$breaks[2]-h$breaks[1])*f(x, -4, 4, 1), lwd=3, lty="dashed", col="mediumpurple")

```

# Exercise 2

Given a discrete probability distribution, defined by the following probabilities:

`0.05, 0.19, 0.14, 0.17, 0.02, 0.11, 0.06, 0.05, 0.04, 0.17`

1. plot the probability density function and the cumulative density function
1. write an algorithm to generate random numbers from the discrete probability distribution

## Solution

**1.** First of all, the sum of the given probabilities is computed to verify it is equal to $1$.

```{r}
p <- c(0.05, 0.19, 0.14, 0.17, 0.02, 0.11, 0.06, 0.05, 0.04, 0.17)
sum(p)
```

Assuming that the discrete $pdf$ is defined in an interval $(1,10)$, the $pdf$ and $cdf$ are plotted by using steps and the `cumsum` function for the $cdf$.

```{r}
options(repr.plot.width=16, repr.plot.height=8)  #to set graph size
par(mfrow=c(1,2),  mar=c(5, 5, 4, 2))            #graphs in the same row

x <- 0:10

plot(0, 0, type='n', xlim=c(0.5,11.5), ylim=c(0,0.2), 
     main="Probability distribution function", cex.main=1.5,
     xlab="x", ylab="f(x)", cex.lab=1.5, las=1)
lines(c(x, 11, 12), c(0, p, 0, 0), type='s', lwd=2.5, col="darkred")

plot(0, 0, type='n', xlim=c(0.5,10.5), ylim=c(0,1), 
     main="Cumulative distribution function", cex.main=1.5,
     xlab="x", ylab="F(x)", cex.lab=1.5, las=1)
lines(c(x, 11), c(0, cumsum(p), 1), type='s', lwd=2.5, col="dodgerblue2")

```

**2.** write an algorithm to generate random numbers from the discrete probability distribution

```{r}
gen_discr <- function(n, x, p){ us <- runif(n, 0, 1)
                                ps <- c(0,p)
                                F <- cumsum(ps)
                                samples <- NULL
                                for (i in 2:length(F)){
                                    s <- which( F[i-1] < us & us <= F[i])
                                    xs <- x[i-1]+ (us[s]−F[i−1])/ps[i]
                                    samples <- c(samples, xs)
                                    }
                                return(samples)
                              }
```

```{r}
x <- 0:10

hist(gen_discr(10000, 0:10, p), freq=FALSE,
     col="lightskyblue", xlab="x", #xlim=c(-6,6), ylim=c(0,0.3),
     main="Sampling from discrete distribution, n=10000")
lines(c(x, 11), c(p, 0, 0), type='s', lwd=3, col="navy")

```

# Exercise 3

Generate random variables from the following distribution:

$$f(x) = \frac{2}{\pi R^2} \sqrt{R^2−x^2}$$

where $−R \leq x \leq R$.

Using the acceptance-rejection algorithm, assume $M= 2/(\pi R)$ and generate $10^4$ random variables, plotting them in an histogram.

## Solution

```{r}
f <- function(x, r){2*sqrt(r^2-x^2)/(pi*r^2)}

acc_rej <- Vectorize(function(n, r){
                y <- NULL
                while(length(y)<n){
                    xs <- runif(n, -r, r)
                    us <- runif(n, 0, 1)
                    M <- 2/(pi*r)
                    ys <- ifelse(us * M < f(xs, r),
                                xs,
                                NA)

                    ys <- na.omit(ys)
                    y <- c(y, ys)
                    }
                return(y)
                })

```

```{r}
n <- 10000
x <- seq(-2,2, 0.05)

options(repr.plot.width=8, repr.plot.height=6)  #to set graph size
par(mar=c(5, 5, 4, 2))                          #graphs in the same row

hist(acc_rej(n, 2), freq=FALSE,
     col="palegreen", xlab="x", cex.lab=1.5, las=1,
     main="Acceptance/Rejection sampling, density histogram")
curve(f(x, 2), col="seagreen4", lt="dotted", lw=3, add=TRUE)


```

<!-- #region -->
# Exercise 4

An important property of the gamma distribution is the so-called _reproductive property_.
Given a sequence of independent random variables $X_j \sim Gamma(\alpha_j, \beta)$, it follows that:

$$ Y =  \sum_{j=1}^n X_j \implies Y \sim Gamma(\alpha,\beta) \mbox{ where } \alpha= \sum_{j=1}^n \alpha_j$$

If $α=m$ is an integer, a random variable from Gamma distribution $Gamma(m,\beta)$ (also known as Erlang distribution) can be obtained by summing $m$ independent exponential random variables $X_j \sim Exp(\beta)$:

$$Y=\beta \sum_{j=1}^n(−ln U_j) =−\beta ln\prod_{j=1}^n U_j $$

Write an algorithm to sample variables from an Erlang distribution $Gamma(m, \beta)$.


## Solution

The Erlang distribution is part of `stats` package in `R`, in particular random data generation is obtained via the `rgamma` function.

This result can be compaired with the one obtained by exploiting the last equation in the text of the exercise, implemented in function `sample_earl`.

<!-- #endregion -->

```{r}
m <- 10
beta <- 0.5
n <- 10000

sample_earl <- function(n, m, beta){ replicate(n, (−1/beta)∗log(prod(runif(m, 0, 1)))) }

options(repr.plot.width=18, repr.plot.height=6)  #to set graph size
par(mfrow=c(1,2), mar=c(5, 5, 4, 2))             #graphs in the same row

hist(sample_earl(n, m, beta), freq=FALSE,
     col="tan1", xlab="x", cex.lab=1.5, las=1,
     main="Sampling from Erlang distribution, manual method")
curve(dgamma(x, m, beta), col="orangered", lt="dotted", lw=3, add=TRUE)

hist(rgamma(n, m, beta), freq=FALSE,
     col="lightpink", xlab="x", cex.lab=1.5, las=1,
     main="Built in R-sampling")
curve(dgamma(x, m, beta), col="hotpink4", lt="dotted", lw=3, add=TRUE)

```

# Exercise 5

One  of  the  first  random  number  generator  was  proposed  by  von  Neumann,  the  so-called _middle square_ algorithm. 

Write `R` code to implement this type of generator and, given a fixed digit number input, square it an remove the leading and trailing digits, in order to return a number with the same number of digits as the original number

_Suggestion_ : after having squared the number, convert it to a list of characters

`(number <- unlist(strsplit(as.character(x.squared),"")))`

and, after having removed the head and tail of the list, convert it back to a number

`(as.numeric(paste(number.after.trimming, collapse="")))`

```{r}
middle_square <- function(n, seed){ 
    x <- NULL
    for (i in 1:n){
        #square the seed
        x.squared <- format(seed^2, scientific=FALSE)
        #convert seed to list of characters
        number <- unlist(strsplit(as.character(x.squared),""))
        #seed length 
        len <- length(unlist(strsplit(as.character(seed),"")))
        #add 0 at the begininning of the number if it does not have the proper number of digits (2*len)
        ifelse( (length(number)<2*len),
                num <- c(rep(0,2*len - length(number)), number),
                num <-  number
               )
        #cut the number head and tail
        trim <- number[(len/2+1):(3*len/2)]
        #convert to number
        gen <- as.numeric(paste(trim, collapse=""))
        x <- c(x, gen)
        seed <- gen
        }
        return(x)
    }
```

```{r}
middle_square(500, 1234430)
```

The algorithm does not work properly. In fact, depending on the seed, the chain of generated number stops.
To try to improve it, a method to control the length of the squared number is implemented. The algorithm adds some $0$s at the beginning of `seed`$^2$ and then, when trimming, eventually shifts the beginning of the sequence in order not to pick a $0$ as beginning digit. 

```{r}
middle_square <- function(n, seed){ 
    x <- NULL
    for (i in 1:n){
        #square the seed
        x.squared <- format(seed^2, scientific=FALSE)
        #convert seed to list of characters
        number <- unlist(strsplit(as.character(x.squared),""))
        #seed length 
        len <- length(unlist(strsplit(as.character(seed),"")))
        #add 0 at the begininning of the number if it does not have the proper number of digits (2*len)
        ifelse( (length(number)<2*len),
                num <- c(rep(0,2*len - length(number)), number),
                num <-  number
               )
        #check if the new number starts with 0 and eventually shift the start of the trimming
        ifelse( number[len/2] == 0,
                trim <- number[(len/2+1):(3*len/2)],
                trim <- number[(len/2):(3*len/2-1)]
               )
        #convert to number
        gen <- as.numeric(paste(trim, collapse=""))
        x <- c(x, gen)
        seed <- gen
        }
        return(x)
    }  
```

```{r}
middle_square(500, 1234430)
```

Despite the applied improvements, the algorithm still does not work properly. In fact, it could happen that there are two zeros close to each other and so that the $0$-check implemented is not enough, hence the numbers generated have not the same number of digits.
To try to avoid this problem, whenever a $0$ is detected as first digit of the trimmed sequence, it is substituted with an arbitrary digit (in this case, a $1$).

```{r}
middle_square <- function(n, seed){ 
    x <- NULL
    for (i in 1:n){
        #square the seed
        x.squared <- format(seed^2, scientific=FALSE)
        #convert seed to list of characters
        number <- unlist(strsplit(as.character(x.squared),""))
        #seed length 
        len <- length(unlist(strsplit(as.character(seed),"")))
        #trimming
        trim <- number[(len/2+1):(3*len/2)]
        #check if the number starts with 0, eventually substitute it with a non-0 digit (randomly chosen)
        ifelse( trim[1]==0,
                trim[1] <- 1,
                trim <- trim               
               )
        #convert to number
        gen <- as.numeric(paste(trim, collapse=""))
        x <- c(x, gen)
        seed <- gen
        }
        return(x)
    }
```

```{r}
middle_square(1000, 1234430)
```

This final algorithm works significantly better that the previous two, but it still has problem with some "unfortunate" seeds, where the sequence is periodic and repeats itself (e.g. 120, 6000...)
