<!-- #region -->
# Exercise 1

The  number  of  particles  emitted  by  a  radioactive  source  during  a  fixed  interval  of  time  ($\Delta t=10$ s) follows a Poisson distribution on the parameter $\mu$.  The number of particles observed during consecutive time intervals is:  4, 1, 3, 1 and 3.

1. Suppose a uniform prior distribution for the parameter $\mu$
    -  determine and draw the posterior distribution for $\mu$, given the data;
    -  evaluate mean, median and variance, both analytically and numerically in R.


2. Suppose a Jeffrey’s prior for the parameter $\mu$
    -  determine and draw the posterior distribution for $\mu$, given the data;
    -  evaluate mean, median and variance, both analytically and numerically in R.


3. Evaluate a $95\%$ credibility interval for the results obtained with both priors. Compare the result with that obtained using a normal approximation for the posterior distribution, with the same mean and standard deviation.
<!-- #endregion -->

## Solution

The Likelihood of several indipendent Poisson processes is given by the product of Poisson distribution:

$$f\left(\left\{x_{j}\right\} | \mu\right)=\prod_{j=1}^{n} f\left(x_{j} | \mu\right) \quad \propto \quad \mu^{\sum x_{j}} \times \mathrm{e}^{(-n \mu)}$$

which is analogue to a Gamma distribution:

$$\begin{array}{c}
\operatorname{Gamma}(x | \alpha, \lambda)=k x^{\alpha-1} \mathrm{e}^{-\lambda x} 
\end{array}$$

with 
$$\begin{array}{c}
k=\frac{\lambda^{\alpha}}{\Gamma(\alpha)} \\
\end{array}$$

### Uniform prior

In the case of a uniform prior $g(\mu)=1 \forall \mu>0$, the Posterior is: 

$$\begin{aligned}
P\left(\mu |\left\{x_{j}\right\}\right) & \propto f\left(\left\{x_{j}\right\} | \mu\right) \times g(\mu) \\
& \propto \mu^{\sum x_{j}} \mathrm{e}^{-n \mu}
\end{aligned}$$

which is exactly a renormalization of the Gamma function defined above, in particular:

$$ \Gamma(\alpha'=\Sigma x_{j}+1, \lambda' = n) = \Gamma(13, 5)$$

```{r}
alpha <- 13
n <- 5
n.sample <- 2000
mu_min <- 0
mu_max <- 10
mu <- seq(mu_min, mu_max, length.out = n.sample)

post_gamma <- dgamma(mu, alpha, n)  

###-----ANALYTICAL SOLUTIONS-----###

mean <- alpha/n
var <- alpha/(n^2)

#The median has no analytical expression

###-----NUMERICAL SOLUTIONS-----###

dmu <- (mu_max-mu_min)/n.sample            

mean_num <- dmu*sum(mu*post_gamma)
median_num <- qgamma(0.5, alpha, n)
var_num <- dmu*sum(mu*mu*post_gamma) - mean_num^2

mu1 <- qgamma(0.025, alpha, n)
mu2 <- qgamma(0.975, alpha, n)

gaus <- dnorm(mu, mean, sqrt(var))

mu1_norm <- mean - 1.96*sqrt(var)
mu2_norm <- mean + 1.96*sqrt(var)

###----------PLOT----------###

options(repr.plot.width=8, repr.plot.height=8)  #to set graph size
par(mar=c(5, 5, 4, 2))

plot(mu, post_gamma, xaxs='i', yaxs='i', type='l', lwd = 3,
     xlim=c(0,10), ylim=c(0,0.6), col = 'brown3',
     main="Posterior with Uniform prior", cex.main=2.5,
     xlab=expression(mu), ylab=expression(paste("P(", mu, " | ",alpha,", n, M)")), 
     cex.lab=2, las=1)
lines(mu, gaus, lwd = 3, lty=2, col='orange')
abline(v=mean, col='grey', lty=2, lwd=2)
abline(v=mu1, col='red', lty=4, lwd=2)
abline(v=mu2, col='green', lty=4, lwd=2)
abline(v=mean_num, col='black', lty=2, lwd=2)
abline(v=mu1_norm, col='blue', lty=4, lwd=2)
abline(v=mu2_norm, col='violetred', lty=4, lwd=2)

legend('topright', col=c("red", "grey", "green",  "blue", "black", "violetred", "orange"), 
       lty=c(4, 2, 4, 4, 2, 4, 2), bty='n', cex=1.5, lwd=3,
       legend = c(parse(text = paste0(' ~mu[1] == ', round(mu1, 4))),
                  parse(text = paste0(' ~bar(mu)[True] == ', round(mean, 4))),
                  parse(text = paste0(' ~mu[2] == ', round(mu2, 4))),
                  parse(text = paste0(' ~mu[1]^norm == ', round(mu1_norm, 4))),
                  parse(text = paste0(' ~bar(mu)[Num] == ', round(mean_num, 4))),
                  parse(text = paste0(' ~mu[2]^norm == ', round(mu2_norm, 4))),
                  "Gaussian approx.")
         )

cat("-----------------------------------\n")
cat("\"True\" values", '\n')
cat("-----------------------------------\n")

cat("Mean=", mean, '\n')
cat("Variance=",var, '\n')
cat("Median cannot be evaluated analytically.\n")

cat("-----------------------------------\n")
cat("Numerical estimation", '\n')
cat("-----------------------------------\n")

cat("Mean=", mean_num, '\n')
cat("Variance=",var_num, '\n')
cat("Median[numerical]=",median_num, '\n')

cat("-----------------------------------\n")
cat("95% Credibility Intervals", '\n')
cat("-----------------------------------\n")
cat("95%C.I. = [", mu1, "; ", mu2,"]", 
    "\n95%C.I.[normal approx.] = [", mu1_norm, "; ", mu2_norm,"]", '\n')


```
### Jeffrey's prior

In the case of a Jeffrey's prior $g(\mu) \propto \frac{1}{\sqrt{\mu}} \forall \mu>0$, the Posterior is: 

$$\begin{aligned}
P\left(\mu |\left\{x_{j}\right\}\right) & \propto f\left(\left\{x_{j}\right\} | \mu\right) \times g(\mu) \\
& \propto \mu^{\sum x_{j}} \mathrm{e}^{-n \mu} \times \frac{1}{\sqrt{\mu}} \\
& \propto \mu^{\sum x_{j}-1 / 2} \mathrm{e}^{-n \mu}
\end{aligned}$$

which is again a renormalization of the Gamma function defined above, in particular:

$$ \Gamma(\alpha'=\Sigma x_{j}+\frac{1}{2}, \lambda'= n) = \Gamma(12.5, 5)$$



```{r}
alpha <- 12.5
n <- 5
n.sample <- 2000
mu_min <- 0
mu_max <- 10
mu <- seq(mu_min, mu_max, length.out = n.sample)

post_gamma <- dgamma(mu, alpha, n)  

###-----ANALYTICAL SOLUTIONS-----###

mean <- alpha/n
var <- alpha/(n^2)

#The median has no analytical expression

###-----NUMERICAL SOLUTIONS-----###

dmu <- (mu_max-mu_min)/n.sample            

mean_num <- dmu*sum(mu*post_gamma)
median_num <- qgamma(0.5, alpha, n)
var_num <- dmu*sum(mu*mu*post_gamma) - mean_num^2

mu1 <- qgamma(0.025, alpha, n)
mu2 <- qgamma(0.975, alpha, n)

gaus <- dnorm(mu, mean, sqrt(var))

mu1_norm <- mean - 1.96*sqrt(var)
mu2_norm <- mean + 1.96*sqrt(var)

###----------PLOT----------###

options(repr.plot.width=8, repr.plot.height=8)  #to set graph size
par(mar=c(5, 5, 4, 2))

plot(mu, post_gamma, xaxs='i', yaxs='i', type='l', lwd = 3,
     xlim=c(0,10), ylim=c(0,0.6), col = 'brown3',
     main="Posterior with Jeffrey's prior", cex.main=2.5,
     xlab=expression(mu), ylab=expression(paste("P(", mu, " | ",alpha,", n, M)")), 
     cex.lab=2, las=1)
lines(mu, gaus, lwd = 3, lty=2, col='orange')
abline(v=mean, col='grey', lty=2, lwd=2)
abline(v=mu1, col='red', lty=4, lwd=2)
abline(v=mu2, col='green', lty=4, lwd=2)
abline(v=mean_num, col='black', lty=2, lwd=2)
abline(v=mu1_norm, col='blue', lty=4, lwd=2)
abline(v=mu2_norm, col='violetred', lty=4, lwd=2)

legend('topright', col=c("red", "grey", "green",  "blue", "black", "violetred", "orange"), 
       lty=c(4, 2, 4, 4, 2, 4, 2), bty='n', cex=1.5, lwd=3,
       legend = c(parse(text = paste0(' ~mu[1] == ', round(mu1, 4))),
                  parse(text = paste0(' ~bar(mu)[True] == ', round(mean, 4))),
                  parse(text = paste0(' ~mu[2] == ', round(mu2, 4))),
                  parse(text = paste0(' ~mu[1]^norm == ', round(mu1_norm, 4))),
                  parse(text = paste0(' ~bar(mu)[Num] == ', round(mean_num, 4))),
                  parse(text = paste0(' ~mu[2]^norm == ', round(mu2_norm, 4))),
                  "Gaussian approx.")
         )

cat("-----------------------------------\n")
cat("\"True\" values", '\n')
cat("-----------------------------------\n")

cat("Mean=", mean, '\n')
cat("Variance=",var, '\n')
cat("Median cannot be evaluated analytically.\n")

cat("-----------------------------------\n")
cat("Numerical estimation", '\n')
cat("-----------------------------------\n")

cat("Mean=", mean_num, '\n')
cat("Variance=",var_num, '\n')
cat("Median[numerical]=",median_num, '\n')

cat("-----------------------------------\n")
cat("95% Credibility Intervals", '\n')
cat("-----------------------------------\n")
cat("95%C.I. = [", mu1, "; ", mu2,"]", 
    "\n95%C.I.[normal approx.] = [", mu1_norm, "; ", mu2_norm,"]", '\n')

```
<!-- #region -->
# Exercise 2

Given the problem of the lightouse, study the case in which both the position along the shore $(\alpha)$ and the distance out at sea $(\beta)$ are unknown.

## Solution

First of all, we assume a uniform distribution as Likelihood for $\theta_k \in [-\frac{\pi}{2},\frac{\pi}{2}]$ (angles of emission):

$$P(\theta_k | \alpha, \beta)=\frac{1}{\pi}$$

which is linked to the parameters $\alpha$ and $\beta$ with the following expression:

$$x_{k} = \alpha+ \beta \tan (\theta_{k})$$.

Applying a change of variables, one obtains a Cauchy distribution for $x_k$ as follows:

$$P\left(x_k | \alpha, \beta\right)=\frac{1}{\pi} \frac{\beta}{\left(x_k-\alpha \right)^{2}+\beta^{2}}$$

which means, as the signals are all indipendent, that the Likelihood $P\left(x  | \alpha, \beta\right)$ will be equal to the product of all the single Likelihoods, i.e.:

$$P\left(x | \alpha, \beta\right)=\prod_k P\left(x_k | \alpha, \beta\right)$$

It can be noted that this is a two parameters problem ($\alpha$ and $\beta$): two prior distributions $P(\alpha)$, $P(\beta)$ are required. Moreover, to solve it, an important assumption is the indipendece of the two distributions, which allow us to write the following expression for Bayes theorem:

$$P\left(\alpha, \beta | x_{k}\right) \propto P\left(x_{k} | \alpha, \beta\right) P(\alpha)P(\beta)$$

Both priors can be assumed as uniform in their ranges $ x \in\left[x_{\min }, x_{\max }\right] $, $y \in\left[0, y_{\max }\right]$:

$$P(\alpha)=\left\{\begin{array}{cl}
\frac{1}{x_{\max }-x_{\min }} & \text { for } \alpha \in\left[x_{\min }, x_{\max }\right] \\
0 & \text { otherwise }
\end{array}\right.$$


$$P(\beta)=\left\{\begin{array}{cl}
\frac{1}{y_{\max }} & \text { for } \beta \in\left[0, y_{\max }\right] \\
0 & \text { otherwise }
\end{array}\right.$$

With this assumptions, the final posterior will be a renormalization (represented by parameter $Z$) of the Likelihood due to Bayes theorem. 

$$\log \left(P\left( \alpha, \beta | x \right) \right)= \log \frac{1}{Z} \prod_k P\left(x_k | \alpha, \beta\right) = \log \frac{1}{Z} \prod_k\frac{1}{\pi} \frac{\beta}{\left(x_k-\alpha \right)^{2}+\beta^{2}} = \text{const} - \sum_k \log( (x_k - \alpha) + \beta^2 )$$

The best parameter estimation is computed taking the maximum of the log-Posterior pdf for both $\alpha$ and $\beta$.
<!-- #endregion -->

```{r}
###----------Generate data----------###

set.seed(2020) 

#number of total data generated
n.sample <- 1000

#grid discretization
n.grid <- 100
#alpha range = [x.min, x.max]
x.min <- -6
x.max <- +6
h <- (x.max - x.min)/n.grid

#beta range = [0, y.max]
y.max <- 5
k <- y.max/n.grid

#grid for alphas and betas
alphas <- seq(from=x.min, by=h, length.out=n.grid)
betas <- seq(from=k, by=k, length.out=n.grid)

#data generation of n angles with uniform distribution, then converted in xs
gen_data <- function(n, alpha, beta) {
    angles <- runif(n, min=-pi/2, max=pi/2)

    xs <- beta * tan(angles) + alpha

    return(xs)
}

#true values of alpha and beta, randomly picked
alpha <- runif(1, x.min, x.max)
beta <- runif(1, 0, y.max)
data <- gen_data(n.sample, alpha, beta)

#select the first n.plot points to represent how the posterior evolves with an increasing number of data
n.str <- readline("Enter data set dimension: ")
n.plot <- as.numeric(unlist(strsplit(n.str, ",")))
dt <- data[1:n.plot]

cat("True position: [", alpha, ",", beta, "]\n")
```

```{r}
#It is easier to work with logarithms and then take the exponential
#The following function computes the log-posterior Cauchy pdf:
#since the prior(alpha) and prior(beta) are uniform, they do not change its shape

#x is our data
#returns a function of alphas given beta
log_posterior <- function (x, beta) {          
  Vectorize(function (alpha) {
              sum( log((beta/pi) / (beta^2 + (x - alpha)^2)) )
           })
}

#create the log-posterior as function of all betas
log_post_b <- log_posterior(dt, betas)      #is a function

#function to make the exponential of the posterior
posterior <- function(as, log_post) {
  log_alphas    <- log_post(as)             # compute log posterior for alpha-grid of values
  log_alpha_max <- max(log_alphas)          # find the max alpha --> estimator for best parameter 
  alphas <- exp(log_alphas - log_alpha_max) # compute posterior by exp(log(post))
}

#to take also beta into account, I create another function to calculate the "2d" log-posterior 
log_post_2d <- function (x_k, alfa, beta) {
  sum( log((beta/pi) / (beta^2 + (x_k - alfa)^2)) )
}

#to combine alpha-beta grid I create a function to calculate the log posterior for both of them
#it is of course vectorized
create_grid <- Vectorize(function(a, b){ log_post_2d(dt, a, b) } )

#outer function takes two vectors and a function and builds a matrix 
#by calling the given function for each combination of the elements in the two vectors 
log_grid <- outer(alphas, betas, create_grid)

#produce a matrix combining the grid values and the previously determined log_post_b
posterior_2d <- matrix(posterior(log_grid, log_post_b), nrow=length(alphas), ncol=length(betas))

#still need to normalize
posterior_2d <- posterior_2d/(k*h*sum(posterior_2d))

```

```{r}
options(repr.plot.width=16, repr.plot.height=7)  #to set graph size
par(mfrow= c(1,2), mar=c(5, 5, 4, 2))

contour(alphas, betas, posterior_2d, las=1, labcex=0.5,
        xlim=c(x.min-0.1,x.max+0.1), ylim=c(-0.1,y.max+0.1),
        nlevels = 6, lwd=2, cex.main= 2.5, cex.lab=2,
        xlab=expression(alpha), ylab=expression(beta), 
        main=expression(paste('p(', alpha, ', ', beta,' | x)')) )
points(alpha, beta, pch=4, col="red", lwd=3)   # true position
legend("topleft", bty='n', c(parse(text = paste0(' "True position, " ~n == ', n.plot))), 
       col="red", pch=4, pt.cex=3, cex=1.5) 

contour(alphas, betas, posterior_2d, las=1, labcex=0.8,
        xlim=c(alpha-1.5, alpha+1.5), ylim=c(beta-1.5,beta+1.5),
        nlevels = 6, lwd=2, cex.main= 2.5, cex.lab=2,
        xlab=expression(alpha), ylab=expression(beta), 
        main=expression(paste('p(', alpha, ', ', beta,' | x) [more zoomed]')))
points(alpha, beta, pch=4, col="red", lwd=3)   # true position
legend("topleft", bty='n', c(parse(text = paste0(' "True position, " ~n == ', n.plot))), 
       col="red", pch=4, pt.cex=3, cex=1.5) 

```

```{r}
options(repr.plot.width=8, repr.plot.height=8)  #to set graph size

persp(alphas, betas , posterior_2d,
      xlab="alpha", ylab="beta", zlab="",
      cex.lab=1.5, lwd=1, border='ivory4',
      xlim=c(x.min,x.max), ylim=c(0,y.max), zlim=c(0,max(posterior_2d)),
      theta=65, phi=45, d=3, col = "aquamarine", shade = 0.3)
title(main=expression(paste('3D p(', alpha, ', ', beta,' | x)')), cex.main=2.5)
```

<!-- #region -->
# Exercise 3

Given the *Signal over Background* example previously discussed, analyze and discuss the following cases:

1. vary the sampling resolution used to generate the data, keeping the same sampling range `xdat <- seq(from=-7*w, to=7*w, by=s*w)`
    - change the resolution $s=\{0.1,0.25,1,2,3\}$;
    - check the effect on the results.
    

2.  Change the ratio $A/B$ used to simulate the data (keeping both positive in accordance with the prior):
    - check the effect on the results


## Solution

### 1. Changing $s$
<!-- #endregion -->

```{r}
# - Generative model
signal <- Vectorize(function(x, a, b, x0, w, t) {
    t * (a*exp(-(x-x0)^2/(2*w^2)) + b)
})

# - Sampling grid for computing posterior
alim    <- c(0.0, 5.0)
blim    <- c(0.2, 2.0)
Nsamp   <- 100
uniGrid <- seq(from=1/(2*Nsamp),to=1-1/(2*Nsamp), by=1/Nsamp)
delta_a <- diff(alim)/Nsamp
delta_b <- diff(blim)/Nsamp
a <- alim[1] + diff(alim)*uniGrid
b <- blim[1] + diff(blim)*uniGrid

# Log posterior
log.post <- function(d, x, a, b, x0, w, t) {
    if(a<0 || b <0) {return(-Inf)} 
    # the effect of the prior
    sum(dpois(d, lambda=signal(x, a, b, x0, w, t), log=TRUE))
}

```

```{r}
# - Define model parameters

x0      <- 0                                         # Signal peak
w       <- 1                                         # Signal width 
s       <- c(0.1, 0.25, 1, 2, 3)                # Change the sampling resolution within the same sampling range
A.true  <- 2                                         # Signal amplitude
B.true  <- 1                                         # Background amplitude
Delta.t <- 5                                         # Exposure time

options(repr.plot.width=16, repr.plot.height=26)     #to set graph size
par(mfrow= c(5,2), mar=c(5, 5, 4, 2))

for (i in 1:length(s)){
    xdat    <- seq(from=-7*w, to=7*w, by=s[i]*w)
    s.true <- signal(xdat, A.true, B.true, x0, w, Delta.t)
    ddat   <- rpois(length(s.true), s.true)
    
    xplot <- seq(from=min(xdat), to=max(xdat), by=0.05*w)
    splot <- signal(xplot, A.true, B.true, x0, w, Delta.t)
    xdat.off <- xdat-(s[i]/2)
    
    ###----------PLOT----------###
    
    plot(xplot, splot, xlab="x", ylab="Signal+Background counts",
         xaxs='i', yaxs='i', type='l', lwd = 3, 
         ylim=c(0, max(ddat, splot)+0.2), #xlim=c(-7.2,7.2), 
         col = 'dodgerblue', main=parse(text = paste0('s == ', s[i])),
         cex.main=2.5, cex.lab=1.5, las=1) 
    lines(xdat.off, ddat, type='s', col='orangered', xlab="", ylab="")
    
    # Compute log unnormalized posterior, z = ln P*(a,b|D), on a regular grid
    z <- matrix(data=NA, nrow=length(a), ncol=length(b))
    fill_matrix <- function(a,b){ log.post(ddat, xdat, a, b, x0, w, Delta.t) }

    z <- outer(a, b, Vectorize(fill_matrix))

    z <- z - max(z)   # set maximum to zero
    norm <- sum(exp(z))
    norm_z <- exp(z)/(delta_a*delta_b*norm)

    # Plot normalized 2D posterior as contours
    
    contour(a, b, norm_z, nlevels = 5, labcex = 0.8, 
            lwd = 2, las=1, cex.main= 2.5, cex.lab=2,
            xlim=c(min(a)-0.1, max(a)+0.1), ylim=c(min(b)-0.1,max(b)+0.1),
            xlab="Amplitude , A", ylab="Background , B",
            main=parse(text = paste0('s == ', s[i])))
    abline(v=2,h=1,col="grey")
    
}
```

Reducing the sampling resolution $s$ (hence, increasing the number of collected data) results in a very narrow posterior distribution, well centered around the "true" values of $A, B$. When we have fewer data, the posterior is wide and not precise.

### 2. Changing $A/B$ ratio

As the important feature is the ratio signal vs. background, I keep the background noise fixed to $B=1$ and change $A$ in a range $[0.1, 0.5, 1, 2, 4]$.

```{r}
# - Define model parameters

x0      <- 0                      # Signal peak
w       <- 1                      # Signal width (chosen within selected values)
A.true  <- c(0.2, 0.5, 1, 2, 4)   # Signal amplitude
B.true  <- 1                      # Background amplitude
Delta.t <- 5                      # Exposure time

options(repr.plot.width=16, repr.plot.height=26)  #to set graph size
par(mfrow= c(5,2), mar=c(5, 5, 4, 2))

for (i in 1:length(A.true)){
    xdat   <- seq(from=-7*w, to=7*w, by=0.5*w)
    s.true <- signal(xdat, A.true[i], B.true, x0, w, Delta.t)
    ddat   <- rpois(length(s.true), s.true)
    
    xplot <- seq(from=min(xdat), to=max(xdat), by=0.05*w)
    splot <- signal(xplot, A.true[i], B.true, x0, w, Delta.t)
    xdat.off <- xdat-0.25
    
    ###----------PLOT----------###
    
    plot(xplot, splot, xlab="x", ylab="Signal+Background counts",
         xaxs='i', yaxs='i', type='l', lwd = 3, 
         xlim=range(xplot), ylim=c(0, max(ddat, splot)+0.2),
         col = 'dodgerblue', main=parse(text = paste0('A/B == ', A.true[i])),
         cex.main=2.5, cex.lab=1.5, las=1) 
    lines(xdat.off, ddat, type='s', col='orangered', xlab="", ylab="")

    # Compute log unnormalized posterior , z = ln P*(a,b|D), on a regular grid
    z <- matrix(data=NA, nrow=length(a), ncol=length(b))
    fill_matrix <- function(a,b){ log.post(ddat, xdat, a, b, x0, w, Delta.t) }

    z <- outer(a, b, Vectorize(fill_matrix))

    z <- z - max(z)   # set maximum to zero
    norm <- sum(exp(z))
    norm_z <- exp(z)/(delta_a*delta_b*norm)

    # Plot normalized 2D posterior as contours
    
    contour(a, b, norm_z, nlevels = 5, labcex = 0.8, 
            lwd = 2, las=1, cex.main= 2.5, cex.lab=2,
            xlim=c(min(a)-0.1, max(a)+0.1), ylim=c(min(b)-0.1,max(b)+0.1),
            xlab="Amplitude , A", ylab="Background , B",
            main=parse(text = paste0('A/B == ', A.true[i])))
    abline(v=A.true[i],h=B.true,col="grey")
    
}
```

Increasing the signal to noise ratio makes the posterior become more centered around the "true" amplitude value, as the noise is less relevant and the signal more significant.
